A fast unified model for parsing and sentence understanding

### 2016-03-16 writing/hard_stack_paper/paper.tex
Second wave of trimming.

\subsection{Shift-reduce parsing}

@@

A shift-reduce parser accepts a sequence of input tokens $\mathbf x = (x_1, \dots, x_N)$ and transitions $\mathbf t = (t_1, \dots, t_T)$, where each $t_i \in \{\shift, \reduce\}$ specifies a \shift~or \reduce~transition (described below). In general a parser may also generate these predictions on-the-fly as it reads the tokens. It reads left-to-right over the input transition sequence, combining the input tokens $\mathbf x$ incrementally into a tree structure. For any binary-branching tree structure over $N$ words, this requires $T = 2N - 1$ transitions.

@@

This subsection describes the core functions and representations of SPINN. We begin with the most basic model, SPINN-NP (for \textit{no parser}), which processes input sentences using parses provided by some external parser.

@@

The buffer is populated with the words in each sentence before the first step of computation. We use word representations from the standard 300-dimensional vector package provided with GloVe \citep{pennington2014glove}. We do not update these representations during training. Instead, we use a learned linear transformation to transform the representation of each input word $\vec{x}$ into a pair of vectors $\langle \vec{h}, \vec{c}\rangle$ that are stored side-by-side in the buffer and used as inputs to the TreeLSTM composition function:
\begin{equation}
\colvec{2}
    {\vec{h}}
    {\vec{c}}
= W_{\text{wd}} \vec{x}
\end{equation}
\paragraph{The composition function}
When a \reduce~operation is performed, vector representations of two tree nodes are popped off of the head of the stack and fed into a {\it composition function}, which is a neural network function that produces a representation for a new tree node that is the parent of the two popped nodes. This new node is then pushed on to the stack.

\paragraph{The contents of the stack and buffer}

@@

This paper evaluates SPINN on a sentence pair classification task. To classify a sentence pair, a feature vector $\vec{x}_{\text{classifier}}$ is first constructed. This feature vector is based on the final representations of each of the two sentences---the representations that appear at the head of the stack for each sentence after the final transition. In particular, the $\vec{h}$ portions of these representations are used. The final feature vector consists of the concatenation of these two vectors, $\vec{h}_{\text{premise}}$ and $\vec{h}_{\text{hypothesis}}$, their difference, and their elementwise product:
\begin{equation}
\vec{x}_{\text{classifier}} = 
\colvec{4}
    {\vec{h}_{\text{premise}}}
    {\vec{h}_{\text{hypothesis}}}
    {\vec{h}_{\text{premise}} - \vec{h}_{\text{hypothesis}}}
    {\vec{h}_{\text{premise}} \odot \vec{h}_{\text{hypothesis}}}
\end{equation}
Following \citet{snli:emnlp2015}, this feature vector is then passed to a series of ReLU neural network layers (the number of layers is a tuned hyperparameter), then passed into a linear transformation, and then finally passed to a softmax layer, which yields a distribution among the three labels. The model is trained in part using a cross-entropy objective over this distribution.

@@

In addition to the stack, the buffer, and the composition function, our full model includes an additional component: the tracking LSTM. The tracking LSTM is a simple low-dimensional sequence-based LSTM RNN that operates in tandem with the model, taking inputs from the buffer and stack at each step. It is meant to maintain a low-resolution summary of the portion of the sentence that has been processed so far, and it is used for two purposes: it supplies feature representations to the transition decision function, which allows the model to stand alone as a parser, and it additionally supplies a secondary input $\vec{e}$ (see Equation~\ref{eqn:lstm1}) to the composition function, allowing context information to leak into the construction of sentence meaning, and forming what is effectively a tree-sequence hybrid model.

The tracking LSTM takes three vectors as inputs at each step (highlighted in yellow in Figure~\ref{fig:m1-views}): the top element of the buffer $\vec{h}_b^1$, which would be moved in a \shift~operation, and the top two elements of the stack $\vec{h}_s^1$ and $\vec{h}_s^2$, which would be composed in a \reduce~operation.

\paragraph{Tracking left context for parsing} The decision function which determines which transitions (and thereby, which parses) SPINN will use needs to take as its input some representation of the current state of the model. We use the hidden state of the tracking LSTM as that input. The tracking LSTM's limited access to the buffer makes this fast to compute, but somewhat more limited than the rich feature sets used in state-of-the-art parsers.

\paragraph{Tracking left context for interpretation} Lexical ambiguity is ubiquitous in natural language. Most words have multiple senses or meanings, and it is generally necessary to use the context in which a word occurs to determine which of its senses or meanings is meant in a given sentence. Simpler sequence-based sentence encoding models like the standard LSTM RNN have an advantage here: when a sequence-based model first processes a word, it has direct access to a state vector that summarizes the left context of that word, which provides some cues for disambiguation. In contrast, when standard tree-structured models first process a word, they only have access to the constituent that the word is merging with, which is often just a single additional word. Feeding a context representation from the tracking LSTM into the composition function is a simple and efficient way to mitigate this disadvantage of tree-structured models.

@@

\todo{[SB] Shorten the previous paras.}

\subsection{Parsing: Predicting transitions from the tracking state}

The model described so far is SPINN-NP (\ii{no parsing}), which uses sequences of \shift~and \reduce~operations from an external parser. To build the fully independent SPINN model, we simply add a single-layer neural network classifier that chooses which operation to perform at each step using the state of the tracking LSTM:
\begin{equation}
\vec{p}_{\text{t}} = \text{softmax}(W_{\text{transition}}\vec{h}_{\text{tracking}})
\end{equation}
The resulting vector $\vec{p}_{\text{t}}$ is a probability distribution over parsing decisions (i.e., predicting  \shift~or \reduce~at the current timestep). In the full SPINN model, we follow whichever transition has a higher predicted probability in this distribution. This prediction module is trained to mimic the decisions of an external parser. In particular, we use the transition sequences corresponding to the constituency parses included in the SNLI corpus.

@@

\vspace{0.5em}
{\noindent\small
{\bf Unlabeled binary parse:} ( ( the cat ) ( sat down ) ) $\Rightarrow$\\
{$\mathbf t$}: \shift, \shift, \reduce, \shift, \shift, \reduce, \reduce\\
{$\mathbf x$}: the, cat, sat, down
}

@@

\begin{quote}\small 
Premise: {\it Girl in a red coat, blue head wrap and jeans is making a snow angel.}

Hypothesis: {\it A girl outside plays in the snow.}

Correct label: {\it entailment}
\end{quote}

@@

\paragraph{Results} Table~\ref{tab:results} shows our results. Our primary evaluation metric for all models is three-way classification accuracy on the SNLI test set. The full SPINN model also generates parsing predictions, and we report a measure of agreement between this model's predictions and the predictions associated with the Stanford PCFG Parser parses included with SNLI (calculated as classification accuracy over \shift~and \reduce~operations). We also report the number of trained parameters in each model and training set accuracy.

@@

\begin{figure}
\resizebox{0.85\linewidth}{!}{
\begin{tikzpicture}
\begin{axis}[xmin=0, ymin=0, xlabel=Batch size, ylabel=Feedforward time (sec),
			 xtick={0,256,512,1024,2048}, legend style={at={(1.1, 0.98)},anchor=north east}]
  \addplot table [x={Batch size}, y=CPU, col sep=tab] {runtime.csv};
  \addlegendentry{CPU \citep{irsoy2014deep}}
  \addplot table [x={Batch size}, y=GPU, col sep=tab] {runtime.csv};
  \addlegendentry{Thin-stack GPU}
\end{axis}
\end{tikzpicture}
}
\caption{Runtime performance results with the thin-stack GPU algorithm.}}
\label{fig:speed}
\end{figure}

>>>>

\subsection{Background: Shift-reduce parsing}

@@

A shift-reduce parser accepts a sequence of input tokens $\mathbf x = (x_1, \dots, x_N)$ and uses transitions $\mathbf t = (t_0, \dots, t_{T-1})$, where each $t_i \in \{\shift, \reduce\}$ specifies a \shift~or \reduce~transition (described below). In general a parser may also generate these predictions on-the-fly as it reads the tokens. It reads left-to-right over the input transition sequence, combining the input tokens $\mathbf x$ incrementally into a tree structure. For any binary-branching tree structure over $N$ words, this requires $T = 2N - 1$ transitions.

@@

\paragraph{The composition function}
When a \reduce~operation is performed, vector representations of two tree nodes are popped off of the head of the stack and fed into a {\it composition function}, which is a neural network function that produces a representation for a new tree node that is the parent of the two popped nodes. This new node is then pushed on to the stack.

\paragraph{The stack and buffer}

@@

We use word representations based on the standard 300-dimensional vector package provided with GloVe \citep{pennington2014glove}. We do not update these representations during training. Instead, we use a learned linear transformation to transform the representation of each input word $\vec{x}_{\text{GloVe}}$ into a pair of vectors $\langle \vec{h}, \vec{c}\rangle$ that are stored side-by-side in the buffer and used as inputs to the TreeLSTM composition function:
\begin{equation}
\colvec{2}
    {\vec{h}}
    {\vec{c}}
= W_{\text{wd}} \vec{x}_{\text{GloVe}}
\end{equation}

@@

This paper evaluates SPINN on a sentence pair classification task. To classify a sentence pair, we first construct a feature vector $\vec{x}_{\text{classifier}}$ based on the SPINN's representation for each of the two sentences---the value at the head of the stacks after the final transition. In particular, the $\vec{h}$ portions of these representations are used. The final feature vector consists of the concatenation of these two vectors, $\vec{h}_{\text{premise}}$ and $\vec{h}_{\text{hypothesis}}$, their difference, and their elementwise product:
\begin{equation}
\vec{x}_{\text{classifier}} = 
\colvec{4}
    {\vec{h}_{\text{premise}}}
    {\vec{h}_{\text{hypothesis}}}
    {\vec{h}_{\text{premise}} - \vec{h}_{\text{hypothesis}}}
    {\vec{h}_{\text{premise}} \odot \vec{h}_{\text{hypothesis}}}
\end{equation}
Following \citet{snli:emnlp2015}, this feature vector is then passed to a series of ReLU neural network layers (or MLP; the number of layers is tuned as a hyperparameter), then passed into a linear transformation, and then finally passed to a softmax layer, which yields a distribution among the three labels.

@@

In addition to the stack, the buffer, and the composition function, our full model includes an additional component: the tracking LSTM. This is a simple low-dimensional sequence-based LSTM RNN that operates in tandem with the model, taking inputs from the buffer and stack at each step. It is meant to maintain a low-resolution summary of the portion of the sentence that has been processed so far, and it is used for two purposes: it supplies feature representations to the transition classifier, which allows the model to stand alone as a parser, and it additionally supplies a secondary input $\vec{e}$ (see Equation~\ref{eqn:lstm1}) to the composition function, allowing context information to leak into the construction of sentence meaning, and forming what is effectively a tree-sequence hybrid model.

The tracking LSTM's inputs (highlighted in yellow in Figure~\ref{fig:m1-views}) are: the top element of the buffer $\vec{h}_b^1$, which would be moved in a \shift~operation, and the top two elements of the stack $\vec{h}_s^1$ and $\vec{h}_s^2$, which would be composed in a \reduce~operation.

\paragraph{Why a tree-sequence hybrid?} Lexical ambiguity is ubiquitous in natural language---most words have multiple senses or meanings, and it is generally necessary to use the context in which a word occurs to determine which of its senses or meanings is meant in a given sentence. Simpler sequence-based sentence encoding models like the standard LSTM RNN have an advantage here: when a sequence-based model first processes a word, it has direct access to a state vector that summarizes the left context of that word, which provides some cues for disambiguation. In contrast, when standard tree-structured models first process a word, they only have access to the constituent that the word is merging with, which is often just a single additional word. Feeding a context representation from the tracking LSTM into the composition function is a simple and efficient way to mitigate this disadvantage of tree-structured models.

@@

\subsection{Parsing: Predicting transitions}

For SPINN to be able to operate on unparsed inputs, it needs to be able to produce its own transition sequence $\mathbf t$ rather than relying on an external parser to supply this as part of the input. To do this, the model predicts $t_t$ at each step using a simple two-way softmax classifier over tracking LSTM features:
\begin{equation}
\vec{p}_{\text{t}} = \text{softmax}(W_{\text{transition}}\vec{h}_{\text{tracking}})
\end{equation}
In the full SPINN model, we follow whichever transition (i.e., \shift~or \reduce) has a higher predicted probability in this distribution. This prediction function is trained to mimic the decisions of an external parser. In particular, we use the transition sequences corresponding to Stanford PCFG Parser constituency parses included with SNLI.

@@

\vspace{0.5em}
{\noindent\small
{\bf Unlabeled binary parse:} ( ( the cat ) ( sat down ) )\\
{$\mathbf t$}: \shift, \shift, \reduce, \shift, \shift, \reduce, \reduce\\
{$\mathbf x$}: the, cat, sat, down
}

@@

\begin{description}\small 
\item[\bf Premise:] {\it Girl in a red coat, blue head wrap and jeans is making a snow angel.}
\item[\bf Hypothesis:] {\it A girl outside plays in the snow.}
\item[\bf Correct label:] entailment
\end{description}

@@

\paragraph{Results} Table~\ref{tab:results} shows our results. Our primary evaluation metric for all models is three-way classification accuracy on the SNLI test set. The full SPINN model also generates parsing predictions, and we report a measure of agreement between this model's predictions and the parses included with SNLI (calculated as classification accuracy over transitions averaged across timesteps). We also report the number of trained parameters in each model and training set accuracy.

@@

\begin{figure}
\resizebox{0.85\linewidth}{!}{
\begin{tikzpicture}
\begin{axis}[xmin=0, ymin=0, xlabel=Batch size, ylabel=Feedforward time (sec),
			 xtick={0,256,512,1024,2048}, legend style={at={(1.1, 0.98)},anchor=north east}]
  \addplot table [x={Batch size}, y=CPU, col sep=tab] {runtime.csv};
  \addlegendentry{CPU \citep{irsoy2014deep}}
  \addplot table [x={Batch size}, y=GPU, col sep=tab] {runtime.csv};
  \addlegendentry{Thin-stack GPU}
\end{axis}
\end{tikzpicture}
}
\caption{Runtime performance results with the thin-stack GPU algorithm.}
\label{fig:speed}
\end{figure}
